Determination of Center of Rotation \\ in Computed Tomography

### 2017-07-21 seminar_cor/seminar_cor_haegele.tex
finished sinogram based

\subsubsection{Cone Beam}
The previously introduced geometries are both 2-dimensional (2D) as they described projections within an x-y-plane onto a one dimensional detector.
To obtain a 3-dimensional (3D) image the scanned object would have to be moved perpendicular to the plane to obtain multiple slice images.
With the 3D cone beam geometry, this is not necessary as it uses a cone shaped beam that projects onto a 2D flat panel detector.
%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{img/conebeam_centralslice.png}
\caption{
A cone beam setup with an object on a rotary table.
It can be seen that only the projection of the central slice is equivalent to a projection in fan beam geometry.
Other projections that are parallel to the s-axis of the detector use samples at multiple heights of the object.
Also the projection of the rotation axis onto the detector is shown which in this case is offset from detectors
z-axis.
Figure from \cite{Yang2011}
}
\label{fig:conebeam}
\end{figure}
%
In \cref{fig:conebeam} it can be seen that there exists a certain line on the detector which will receive the same input as a fan beam detector.
All other lines read projections through multiple planes of the object and are thus not rebinnable to parallel beam.

@@

\subsubsection{Cross-Correlation}
Yang et al.~\cite{Yang2011} proposed another sinogram based method for COR determination using image cross-correllation.
Given a sinogram in a prallel beam geometry over $360^\circ$ (which can also be obtained by rebinning a fan-beam sinogram), the sinogram is divided in half so that both halfs cover a $180^\circ$ range.
As the information in both sub sinograms is redundant in that way that they are mirror images of each other,
flipping one of them will result in two sinograms over $180^\circ$ with identical information.
%
\begin{equation}
R(r,\gamma)^{[0,\pi]} = R(-r,\gamma)^{[\pi,2\pi]}
\end{equation}
%
A shift of COR, as explained in the beginning of \cref{sec:cordetermination}, results in a shift of the sinogram along the $r$-axis.
Thus, the two sub sinograms are identical under this shift.
%
\begin{equation}
R_\tau(r+\tau,\gamma)^{[0,\pi]} = R_\tau(-(r+\tau),\gamma)^{[\pi,2\pi]}
\end{equation}
%
Using the 1D the cross correlation of the two sinograms along the $r$-axis, the shift can be determined as follows.
%
\begin{equation}
2\tau = \underset{r}{arg\,max}\quad R_\tau(r,\gamma)^{[0,\pi]} \star R_\tau(-r,\gamma)^{[\pi,2\pi]}
\end{equation}
%

>>>>

\subsubsection{Cone Beam}\label{sssec:conebeam}
The previously introduced geometries are both 2-dimensional (2D) as they described projections within an x-y-plane onto a one dimensional detector.
To obtain a 3-dimensional (3D) image the scanned object would have to be moved perpendicular to the plane to obtain multiple slice images.
With the 3D cone beam geometry, this is not necessary as it uses a cone shaped beam that projects onto a 2D flat panel detector.
%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{img/conebeam_centralslice.png}
\caption{
A cone beam setup with an object on a rotary table.
It can be seen that only the projection of the central slice is equivalent to a projection in fan beam geometry.
Other projections that are parallel to the s-axis of the detector use samples at multiple heights of the object.
Also the projection of the rotation axis onto the detector is shown which in this case is offset from detectors
z-axis.
Figure from \cite{Yang2011}
}
\label{fig:conebeam}
\end{figure}
%
In \cref{fig:conebeam} it can be seen that there exists a certain line on the detector which will receive the same input as a fan beam detector.
All other lines read projections through multiple planes of the object and are thus not rebinnable to parallel beam.

@@

\subsubsection{Cross-Correlation}
Yang et al.~\cite{Yang2011} proposed another sinogram based method for COR determination using image cross-correllation.
Given a sinogram in a prallel beam geometry over $360^\circ$ (which can also be obtained by rebinning a fan-beam sinogram), the sinogram is divided in half so that both halfs cover a $180^\circ$ range.
As the information in both sub sinograms is redundant in that way that they are mirror images of each other,
flipping one of them will result in two sinograms over $180^\circ$ with identical information.
%
\begin{equation}
R(r,\gamma)^{[0,\pi]} = R(-r,\gamma)^{[\pi,2\pi]}
\end{equation}
%
A shift of COR, as explained in the beginning of \cref{sec:cordetermination}, results in a shift of the sinogram along the $r$-axis (see~\cref{eq:radonoffset}).
Thus, the two sub sinograms are identical under this shift.
%
\begin{equation}
R_\tau(r+\tau,\gamma)^{[0,\pi]} = R_\tau(-(r+\tau),\gamma)^{[\pi,2\pi]}
\end{equation}
%
Using the 1D cross correlation of the two sinograms along the $r$-axis, the shift can be determined as follows.
%
\begin{equation}
2\tau = \underset{r}{arg\,max}\quad R_\tau(r,\gamma)^{[0,\pi]} \star R_\tau(-r,\gamma)^{[\pi,2\pi]}
\end{equation}
%
The authors~\cite{Yang2011} provided evidence that their algorithm is robust against Gaussian noise even for large variance values.
They also stated that their algorithm can be used for 3D CT with cone beam geometry when using the central slice (see \cref{sssec:conebeam}).
As locating the central slice is not trivial and depends on calibration as well, it was shown that the algorithm still performs well for sinograms that slightly deviate from that of the central slice.
